\ \\ \LARGE\bf Speculative Evaluation in Particle Swarm Optimization \thanks{Matthew Gardner, Andrew McNabb, and Kevin Seppi are with the Department of Computer Science, Brigham Young University, 3361 TMCB, Provo, UT 84602 (phone: 801-422-8717; email: \{mjg82,a,k\}@cs.byu.edu).}

### 2010-03-30 honors_thesis/journal_paper/paper.tex
Fixed some broken references

We first define a helper function to update the personal and neighborhood bests, $\pbest$ and $\nbest$.
This algorithm is shown in~\alg{alg:bests}.

@@

\begin{algorithm} [h!]
  \caption{Standard PSO (minimizing)}
  \label{alg:standard}
  \begin{algorithmic}[1]
    \STATE t=0
    \STATE Initialize the swarm
%    \FOR{$i \in \swarm$}
%      \STATE {$\ppos_{i,t} \leftarrow Random position$}
%      \STATE {$\pvel_{i,t} \leftarrow Random velocities$}
%      \STATE {$\fpbest_{i,t} \leftarrow Max Float$}
%      \STATE {$\fnbest_{i,t} \leftarrow Max Float$}
%    \ENDFOR
    \REPEAT
        \STATE $t=t+1$
        \FOR{each particle in the swarm}
          \STATE {$\fppos_{t} \leftarrow \ofunc(\ppos_{t})$}
        \ENDFOR
        \FOR{each particle in the swarm}
          \STATE $\nURand_{t} \Leftarrow RandomVector()$
          \STATE $\nURand_{t} \Leftarrow RandomVector()$
          \STATE UpdateBests
          \STATE Update $\pvel_{t+1}$ as shown in equation~\eqref{eq:velupdate} 
          \STATE Update $\ppos_{t+1}$ as shown in equation~\eqref{eq:posupdate}
        \ENDFOR
    \UNTIL {$t > max \ iterations$ or convergence criteria met}
  \end{algorithmic}
\end{algorithm}

Using this helper function we can write PSO as shown in~\alg{alg:standard}.
This algorithm can be transformed, while retaining it exact behavior, by
``unrolling'' one iteration of the ``Repeat'' loop, redundantly expressing work done in the ``if'''s
in~\alg{alg:bests}
using the indicator functions $\indic_{t+1}$ and re-writing the computation of the next velocity,
positions and objective function value using 
equations~\eqref{eq:vel2update}, \eqref{eq:pos2update} and \eqref{eq:val2update} 
as shown in the new version of the algorithm as shown in~\alg{alg:unrolled}

\begin{algorithm} [h!]
  \caption{Unrolled PSO}
  \label{alg:unrolled}
  \begin{algorithmic}[1]
    \STATE t=0
    \STATE Initialize the swarm including
%    \FOR{$i \in \swarm$}
%      \STATE {$\ppos_{i,t} \leftarrow Random position$}
%      \STATE {$\pvel_{i,t} \leftarrow Random velocities$}
%      \STATE {$\fpbest_{i,t} \leftarrow Max Float$}
%      \STATE {$\fnbest_{i,t} \leftarrow Max Float$}
%    \ENDFOR
    \REPEAT
        \STATE $t=t+1$
        \FOR{each particle in the swarm}
          \label{alg2:fatitert}
          \STATE DISPATCH the computation of {$\ofunc(\ppos_{t})$} to a seperate processor
        \ENDFOR
        \FOR{each particle in the swarm}
          \STATE $\nURand_{t} \Leftarrow RandomVector()$
          \STATE $\nURand_{t} \Leftarrow RandomVector()$
          \FOR{$c \in \caseset$}
            \STATE Compute $\specvel_{t+1}^c$ and $\specpos_{t+1}^c$.
	    Examples shown in equations~\eqref{eq:defvcasexn} and~\eqref{eq:defpcasexn}
            \label{alg2:fatitertp1}
            \STATE DISPATCH computation of $\ofunc(\specpos_{t+1}^c)$ to a seperate processor.
	    Example shown in equation~\eqref{eq:deffcasexn}.
	    // Note that NONE of the computations, $\specvel_{t+1}^c$,
	    $\specpos_{t+1}^c$ and $\ofunc(\specpos_{t+1}^c)$ depend on $\ofunc(\ppos_{t})$.
          \ENDFOR
        \ENDFOR
        \STATE WAIT for all function evaluations dispatched above to complete
        \FOR{each particle in the swarm}
          \FOR{$c \in \caseset$}
            \STATE Compute $\indic_{t+1}^c$
	           Examples shown in equation~\eqref{eq:deficasexn} and~\eqref{eq:defpcasexn}
	  \ENDFOR 
            \label{alg2:startcut}
          \STATE Compute $\pvel_{t+1}$ as shown in equation~\eqref{eq:vel2update} 
          \STATE Compute $\ppos_{t+1}$ as shown in equation~\eqref{eq:pos2update}
          \STATE Compute $\ofunc(ppos_{t+1})$ as shown in equation~\eqref{eq:val2update}
	  \STATE UpdateBests
            \label{alg2:endcut}
        \ENDFOR

@@

We outline three main improvements to speculative evaluation.  First, in
\sec{pickbest} we describe a method that uses all of the information found in
doing speculative evaluations.  Then Sections~\ref{sec:pruning}
through~\ref{sec:wrong} present a technique that reduces the number of
speculative evaluations that need to be done for each particle.  Finally,
\sec{manyiters} shows a method for speculating several iterations ahead,
instead of just one.  \sec{other} presents a few other ideas that seem
promising but need further development.

>>>>

We first define a helper function to update the personal and neighborhood bests, $\pbest$ and $\nbest$.
This algorithm is shown in~\alg{bests}.

@@

\begin{algorithm} [ht!]
  \caption{Standard PSO (minimizing)}
  \label{alg:standard}
  \begin{algorithmic}[1]
    \STATE t=0
    \STATE Initialize the swarm
%    \FOR{$i \in \swarm$}
%      \STATE {$\ppos_{i,t} \leftarrow Random position$}
%      \STATE {$\pvel_{i,t} \leftarrow Random velocities$}
%      \STATE {$\fpbest_{i,t} \leftarrow Max Float$}
%      \STATE {$\fnbest_{i,t} \leftarrow Max Float$}
%    \ENDFOR
    \REPEAT
        \STATE $t=t+1$
        \FOR{each particle in the swarm}
          \STATE {$\fppos_{t} \leftarrow \ofunc(\ppos_{t})$}
        \ENDFOR
        \FOR{each particle in the swarm}
          \STATE $\nURand_{t} \Leftarrow RandomVector()$
          \STATE $\nURand_{t} \Leftarrow RandomVector()$
          \STATE UpdateBests
          \STATE Update $\pvel_{t+1}$ as shown in equation~\eqref{eq:velupdate} 
          \STATE Update $\ppos_{t+1}$ as shown in equation~\eqref{eq:posupdate}
        \ENDFOR
    \UNTIL {$t > max \ iterations$ or convergence criteria met}
  \end{algorithmic}
\end{algorithm}

Using this helper function we can write PSO as shown in~\alg{standard}.
This algorithm can be transformed, while retaining it exact behavior, by
``unrolling'' one iteration of the ``Repeat'' loop, redundantly expressing work done in the ``if'''s
in~\alg{bests}
using the indicator functions $\indic_{t+1}$ and re-writing the computation of the next velocity,
positions and objective function value using 
equations~\eqref{eq:vel2update}, \eqref{eq:pos2update} and \eqref{eq:val2update} 
as shown in the new version of the algorithm as shown in~\alg{unrolled}

\begin{algorithm} [ht!]
  \caption{Unrolled PSO}
  \label{alg:unrolled}
  \begin{algorithmic}[1]
    \STATE t=0
    \STATE Initialize the swarm including
%    \FOR{$i \in \swarm$}
%      \STATE {$\ppos_{i,t} \leftarrow Random position$}
%      \STATE {$\pvel_{i,t} \leftarrow Random velocities$}
%      \STATE {$\fpbest_{i,t} \leftarrow Max Float$}
%      \STATE {$\fnbest_{i,t} \leftarrow Max Float$}
%    \ENDFOR
    \REPEAT
        \STATE $t=t+1$
        \FOR{each particle in the swarm}
          \label{alg2:fatitert}
          \STATE DISPATCH the computation of {$\ofunc(\ppos_{t})$} to a seperate processor
        \ENDFOR
        \FOR{each particle in the swarm}
          \STATE $\nURand_{t} \Leftarrow RandomVector()$
          \STATE $\nURand_{t} \Leftarrow RandomVector()$
          \FOR{$c \in \caseset$}
            \STATE Compute $\specvel_{t+1}^c$ and $\specpos_{t+1}^c$.
	    Examples shown in equations~\eqref{eq:defvcasexn} and~\eqref{eq:defpcasexn}
            \label{alg2:fatitertp1}
            \STATE DISPATCH computation of $\ofunc(\specpos_{t+1}^c)$ to a seperate processor.
	    Example shown in equation~\eqref{eq:deffcasexn}.
	    // Note that NONE of the computations, $\specvel_{t+1}^c$,
	    $\specpos_{t+1}^c$ and $\ofunc(\specpos_{t+1}^c)$ depend on $\ofunc(\ppos_{t})$.
          \ENDFOR
        \ENDFOR
        \STATE WAIT for all function evaluations dispatched above to complete
        \FOR{each particle in the swarm}
          \FOR{$c \in \caseset$}
            \STATE Compute $\indic_{t+1}^c$
	           Examples shown in equation~\eqref{eq:deficasexn} and~\eqref{eq:defpcasexn}
	  \ENDFOR 
            \label{alg2:startcut}
          \STATE Compute $\pvel_{t+1}$ as shown in equation~\eqref{eq:vel2update} 
          \STATE Compute $\ppos_{t+1}$ as shown in equation~\eqref{eq:pos2update}
          \STATE Compute $\ofunc(ppos_{t+1})$ as shown in equation~\eqref{eq:val2update}
	  \STATE UpdateBests
            \label{alg2:endcut}
        \ENDFOR

@@

We outline three main improvements to speculative evaluation.  First, in
\sec{pickbest} we describe a method that uses all of the information found in
doing speculative evaluations.  Then Sections~\ref{sec:pruning}
through~\ref{sec:wrong} present a technique that reduces the number of
speculative evaluations that need to be done for each particle.  Finally,
\sec{manyiters} shows a method for speculating several iterations ahead,
instead of just one.
